{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-03-10 thesis/landslide.tex
decribe icb, bpor

First of all, note that Algorithm~\ref{alg:icb} is structured in a way that repeats interleavings
with fewer than $n$ preemptions that have already been checked in previous iterations of the outer loop.
This is because the number of preemptions in each branch is not known in advance;
rather, the state space must always be explored in an overall depth-first approach,
at best skipping too-preemptful interleavings as they are encountered.
As simple as it would be to state ``{\bf foreach} $S \in {\mathsf{sort}_P(U)}$'' in pseudocode,
implementing such an ordering would be much less straightforward. %
% TODO: post committee review
%\footnote{Translating conference papers' pseudocode into usable implementations is often quite difficult \cite{itg2}.}
The core ICB loop appears in {\tt time\_travel()} in {\tt landslide.c}.
%(although not explicitly structured as a C-style loop in the code)

@@

Secondly, prior work \cite{bpor} has shown that
when combined with DPOR to prune equivalent interleavings,
DPOR's reduction might not be sound with respect to the subset of $U$ under $P \le n$.
That work introduced Bounded Partial Order Reduction (BPOR),
a compatibility extension to DPOR for ICB to address this problem.

To summarize, when DPOR identifies some interleaving $I_{ij}$ to test,
it may not be possible to execute $T(t_j)$ after $t_i$ without exceeding the current preemption bound.
However, there may exist another interleaving $J_{i'j}$ within the bound which runs $T(t_j)$ before $t_i$.
If a DPOR implementation na\"ively configured with ICB simply skipped $I$ on account of the preemption bound,
there's no guarantee $J$ would get marked for exploration
from any other iteration and/or any other pair of conflicting transitions.
Even though restricting the state space to a certain maximum preemption bound is already unsound
in terms of losing full concurrency coverage,
failing to test even $J$ would be a failure of DPOR itself
to soundly prune the already-reduced state space defined by the bound.
Hence the need for BPOR, to ensure that if such an alternative $J$ to $I$ exists within the bound,
$J$ gets marked for exploration immediately.

%I refer the reader interested in the algorithm's details to the citation;
%this section will discuss Landslide's implementation.
% TODO: say how u do it

>>>>

\subsubsection{Implementation}

First of all, note that Algorithm~\ref{alg:icb} is structured in a way that repeats interleavings
with fewer than $n$ preemptions that have already been checked in previous iterations of the outer loop.
This is because the number of preemptions in each branch is not known in advance;
rather, the state space must always be explored in an overall depth-first approach,
at best skipping too-preemptful interleavings as they are encountered.
As simple as it would be to state ``{\bf foreach} $S \in {\mathsf{sort}_P(U)}$'' in pseudocode,
implementing such an ordering would be much less straightforward. %
% TODO: post committee review
%\footnote{Translating conference papers' pseudocode into usable implementations is often quite difficult \cite{itg2}.}

Therefore, Landslide's ICB implementation combines with DPOR
when tagging new branches to explore at the end of each branch:
just as DPOR skips alternate interleavings that are memory-independent,
ICB further filters interleavings requiring more preemptions than the current bound
out of the to-explore set.
The macro {\tt ICB\_BLOCKED},
defined in {\tt schedule.h},
decides if a given thread would require a preemption beyond the current bound to switch to.
The DPOR implementation then checks, for some $I_{ij}$ it wants to mark for exploration,
whether {\tt ICB\_BLOCKED(}$T_j${\tt )} at the state after $t_i$,
and skips it if so ({\tt tag\_good\_sibling()}/{\tt tag\_all\_siblings()}).
\footnote{Since ``voluntary'' context switches (e.g. arising from {\tt yield()})
are often necessary for correct execution,
{\tt ICB\_BLOCKED} does not count such switches towards the preemption count.
Therefore, within a certain preemption bound $b$,
interleavings with more than $b$ context switches may still be tested.}

Then, the entire state space is repeated with increasing bound until no such are filtered.
This core ICB loop appears in {\tt time\_travel()} in {\tt landslide.c}.
Although not explicitly structured as a C-style loop in the code,
it resets Landslide's progress through the state space,
allowing exploration to continue until it finally observes all interleavings to have fewer preemptions than the bound.

@@

Prior work \cite{bpor} has shown that
when combined with DPOR to prune equivalent interleavings,
DPOR's reduction might not be sound with respect to the subset of $U$ under $P \le n$.
That work introduced Bounded Partial Order Reduction (BPOR),
a compatibility extension to DPOR for ICB to address this problem.

To summarize, when DPOR identifies some interleaving $I_{ij}$ to test,
it may not be possible to execute $T(t_j)$ after $t_i$ without exceeding the current preemption bound.
However, there may exist another interleaving $J_{i'j}$ within the bound which runs $T(t_j)$ before $t_i$.
If a DPOR implementation na\"ively configured with ICB simply skipped $I$ on account of the preemption bound,
$J$ may not get marked for exploration
from any other iteration and/or pair of conflicting transitions.
Even though restricting the state space to a certain maximum preemption bound is already unsound
in terms of losing full interleaving coverage,
failing to test even $J$ would be a failure of DPOR itself
to soundly prune the already-reduced state space defined by that bound.
Hence the need for BPOR, to ensure that if such an alternative $J$ to $I$ exists within the bound,
it gets marked for exploration immediately.

To implement BPOR,
whenever {\tt ICB\_BLOCKED} causes DPOR to skip an interleaving,
Landslide searches all transitions $t_k \in S$
such that $t_k \prec_S t_i$ and $T(t_k) = T(t_i)$ and
$\neg$\{$\exists t_l \in S$ such that $t_k \preceq_S t_l \prec_S t_i$ and $T(t_l) = T_j$\}
({\tt stop\_bpor\_backtracking()}).
All $I_{kj}$s which can be tested within the preemption bound are marked instead of $I_{ij}$
({\tt tag\_reachable\_aunts()}).
The reader interested in further algorithm details and the corresponding soundness proof is referred to \cite{bpor}.
